---
title: "BioDiversity"
author: "Calla Bush St George"
date: "`r Sys.Date()`"
output:
  html_document: 
    code_folding: show
    theme: spacelab
    highlight: pygments
    keep_md: yes
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
      toc_depth: 3
  keep_md: true  
editor_options: 
  chunk_output_type: console
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE,
                      fig.align = "center",
                      fig.path = "../figures/04_Biodiversity/",
                      dev = "png", dpi = 200) 

# send any figure output to this folder 

```

# Setting the Environment 

### Set my seed
```{r set seed}
# Any number can be chose
set.seed(567890)
```

## Load Libraries 
```{r load-packages}

pacman::p_load(tidyverse, devtools, patchwork, iNEXT, phyloseq,rstatix, ggpubr,
               install = FALSE)
```

## Load in Data 
```{r load-data}
load("data/02_PreProcessing/raw_preprocessed_physeq.RData")
raw_preprocessed_physeq

# Intuition Check 
min(sample_sums(raw_preprocessed_physeq))


# Setting colors for gut sections 
host_species_colors <- c(
  "Acanthurus nigros" = "dodgerblue4",
  "Acanthurus achilles" = "#FF5733",
  "Acanthurus nigrofuscus" = "olivedrab",
  "Acanthurus olivaceus" = "purple4",
  "Acanthurus triostegus" = "maroon2",
  "Ctenochaetus striatus" = "yellow2",
  "Naso brevirostris" = "red2",
  "Naso lituratus" = "royalblue",
  "Naso tonganus" = "salmon2",
  "Odax pullus" = "springgreen2",
  "Zebrasoma flavescens" = "mediumorchid",
  "Zebrasoma scopas" = "magenta",
  "Zebrasoma velifer" = "goldenrod")

# Make metadata dataframe
metadata_df <-
  raw_preprocessed_physeq %>%
  sample_data() %>%
  data.frame()

```



# Goals

1. Calculate the Hill Diversity of the samples. 
2. Evaluate the rarefaction curves. 
3. Evaluate the Diversity values. 
4. Makes notes of specific samples and their seq depth. 

# Diversity Calculations with iNEXT 

```{r calc-div}
# prepare input data 
iNEXT_input_df <- 
  raw_preprocessed_physeq %>%
  otu_table() %>%
  data.frame()
# Quick check
dim(iNEXT_input_df)

# Run iNEXT: Calculate the Hill Numbers 
# Note that: Species in ROWS, Samples in COLUMNS 
# Remember to set the seed! 
#iNEXT_data <- iNEXT(iNEXT_input_df, 
 #                  q = c(0,1,2), datatype = "abundance")

# Save the file
#save(iNEXT_data, file = "data/04_Biodiversity/iNEXT_data.RData")
```

# Evaluate the Diversity! 
```{r load-div}
load("data/04_Biodiversity/iNEXT_data.RData")
str(iNEXT_data)
typeof(iNEXT_data)
```

# Plot Diversity 
```{r plot-rarefaction}
# Prepare Colors 
color_df <- 
  iNEXT_input_df %>%
  colnames() %>%
  data.frame()
# Check
head(color_df)
# Rename the column 
colnames(color_df)[1] <- "names"
# Check
head(color_df)


# Make a helper dataframe for plotting with colors 
iNEXT_color_df <- 
  color_df %>%
  # Fix the names for merging
  mutate(names = gsub(names, pattern = "[.]", replace = "-"),
         names = gsub(names, pattern = "X",  replace = "")) %>%
  # Merge with metadata
  left_join(metadata_df, by = "names") %>%
   #Merge with colors for plotting with ggiNEXT
  left_join(y=data.frame(host_species_colors = host_species_colors,
            host_species = names(host_species_colors)),
            by = "host_species")
```

# Plot Rarefaction with `ggiNEXT`

```{r ggiNEXT, fig.width=8, fig.height=3.5}
# Plot rarefaction! 
# rarefaction/extrapolation curve, type = 1 

# Order q: 
  # 0 = Richness/ Number of Total taxa
  # 1 = Exponential Shannon / Number of "Common" taxa
  # 2 = Inverse Simpson / Number of "Dominant" taxa 

#ggiNEXT(iNEXT_data, type = 1, facet.var = "Order.q") + 
#  facet_wrap(~Order.q, scales = "fixed") + 
#  scale_color_manual(values = iNEXT_color_df$host_species_colors, guide = FALSE) + 
 # scale_fill_manual(values = iNEXT_color_df$host_specie_colors, guide = FALSE) #+ 
 # scale_shape_manual(values = base::rep(17, nsamples(raw_preprocessed_physeq)),
   #                  guide = FALSE) +
 # theme(legend.position = "none")
```


# Manually plot Diversity 

## Rarefaction
```{r iNEXT-manual}

iNEXT_manual_df <- 
  iNEXT_data$iNextEst$size_based %>%
  dplyr::rename(names = Assemblage) %>%
  # Fix the samples names 
  mutate(names = gsub(names, pattern = "[.]", replace = "-"),
         names = gsub(names, pattern = "X", replace = "")) %>%
  # join with metadata 
  left_join(., metadata_df, by = "names") %>%
  # Add colors to data frame
  left_join(., data.frame(host_species_colors = host_species_colors,
                          host_species = names(host_species_colors)),
            by = "host_species") 

# Inspect 
dim(iNEXT_manual_df)
str(iNEXT_manual_df)

# Plot it - Rarefaction Curve 
iNEXT_manual_df %>%
  # Filter out rows that are calcaulted by rarefaction from iNEXT
  dplyr::filter(Method == "Extrapolation") %>%
  # Make the actual rarefaction plot with 
  # the # of sequences on the x-axis and diversity on the y-axis
  # You can choose to pick one diversity value or plot all three 
  ggplot(aes(x = m, y= qD, color = host_species, group = names)) + 
  # line 
  geom_line() + 
  #geom_point() + 
  # Challenge: Facet with gut section
  facet_grid(Order.q~host_species, scales = "fixed") + 
  scale_color_manual(values = host_species_colors) + 
  theme(legend.position = "bottom")
```


# Diversity vs Gut Section 

```{r div-vs-, fig.height=3.5, fig.width=6}

iNEXT_manual_df %>%
  dplyr::filter(Method == "Observed") %>%
  ggplot(aes(x = host_species, y = qD)) + 
  geom_boxplot(alpha = 0.5, outlier.shape = NA) +
  facet_wrap(.~Order.q, scales = "free") + 
  geom_point(aes(color = host_species)) + 
  stat_smooth() + 
  labs(x = "Host species", y = "# of ASVs") + 
  scale_color_manual(values = host_species_colors) + 
  theme(legend.position = "bottom")

```


# Session Information 
```{r session-info}
# Ensure reproducibility 
devtools::session_info()
```
